Dirac neutrino mass generation from Majorana messenger

### 2019-09-17 draft.tex
Updates from Overleaf

\begin{abstract}    
We propose a model for one-loop Dirac neutrino masses with Majorana mediators which can be dark matter candidates if they are the lightest states circulating the loop. This model is restricted by  neutrino physics,  lepton flavor violation and cosmological constrains, including  
the effective number light neutrinos $N_{\text{eff}}$. Here we find that $M_{Z^{\prime}}/g^{\prime} \gtrsim 60\ \text{TeV}$ in order to satisfy the conditions of $N_{\text{eff}}$..
\end{abstract}

@@

\section{Beyond the minimal model: cosmological constraints}
\label{sec:CosmoConstraints}
It turns that the $Z'$ portal also allows to probe the model through modifications on the cosmological history of the Universe, namely, via additional contributions to the effective number of relativistic degrees of freedom $N_{\text{eff}}$~\cite{Dolgov:2002wy}. In this model, these contributions arise from the presence of the right-handed neutrinos and may be expected to be sizeable precisely due to the large $\operatorname{U}(1)_{\mathcal{D}}$~charges of the $\nu_R$'s.  Nevertheless, since the right-handed neutrinos do not couple directly to the rest of the SM particles (see Eq.~(\ref{eq:LagY}))  they  decouple early enough from the thermal bath and, therefore, do not modify the SM prediction for $N_{\text{eff}}$.

@@

The results for $\Delta N_{\text{eff}}$ as a function of $M_{Z^{\prime}}/g^{\prime}$ are displayed in Fig.~\ref{fig:Neff}. It follows that for small rartios $M_{Z'}/g'$  the variation in the number of relativistic degrees of freedom of the new species is large ($\Delta N_{\text{eff}}\gtrsim0.3$ corresponds to decoupling temperatures $T^{\nu_R}_{\text{dec}}\lesssim2$ GeV) and vice versa.
% The bound that come from Planck~\cite{Aghanim:2018eyx} is stronger than the bound of LEP~\cite{Heeck:2014zfa}. and even LHC, see below
It follows that for the $\operatorname{U}(1)_{G}$ model the Planck+BAO upper bound implies that $M_{Z^{\prime}}/g^{\prime} \gtrsim 40$ TeV while for the $\operatorname{U}(1)_{R}$ model a more stringent lower bound applies $M_{Z^{\prime}}/g^{\prime} \gtrsim 60$ TeV. 
%in order to satisfy the conditions of $N_{\text{eff}}$ for all the symmetries studied here.
On the other hand, it is remarkable the fact the future CMB experiments~\cite{Benson:2014qhw,Abitbol:2019nhf} have the potential to entirely probe all the not-so minimal models~\cite{Abazajian:2019oqj}.

%
\begin{figure}
\centering
\includegraphics[scale=1]{D_Neff.pdf}
\caption{Contribution to the number of extra relativistic degrees of freedom ($\Delta N_{\text{eff}}$) in function of $M_{Z^{\prime}}/g^{\prime}$. The region above the gray line is excluded by the measurements at $2\sigma$ from PLANCK~\cite{Aghanim:2018eyx}. The horizontal dashed lines show the projected sensitivity of the future experiments SPT-3G/SO~\cite{Benson:2014qhw} and CMB-S4~\cite{Abitbol:2019nhf}. }
\label{fig:Neff}
\end{figure}

%Prospect on $N_{\text{eff}}$~\cite{Abazajian:2019oqj}

\section{Collider constraints}
The recasting of the latest ATLAS results for the search of dilepton resonances using $139\ \text{fb}^{-1}$~\cite{Aad:2019fac} was done
in~\cite{Chiang:2019ajm} for the $\operatorname{U}(1)_{B-L}$ model.
The green (upper) region in fig.~\ref{fig:u1blc} shows the excluded region at $95\%\ \text{C.L.}$ To ease the comparison with other results, we show the exclusion as function of $M_{Z'}/g^{\prime}$. In particular, the limit from LEP for $\operatorname{U}(1)_{B-L}$ model is~\cite{Carena:2004xs,Heeck:2014zfa}\footnote{The constraint as function of $h=-1-l$ is given in~\cite{Okada:2016tci}. }, 
\begin{align}
  M_{Z'}/g^{\prime}>6.7\ \text{TeV}\,,&&\text{at $95\%$ C.L }\,,
\end{align}
which is obtained from the search for effective four-lepton operators and is valid for $M_{Z'}\gg 200\ \text{GeV}$. This constraint corresponds to the excluded magenta (lower) region of fig.~\ref{fig:u1blc} and start to be relevant for $M_{Z'}>5.8\ \text{TeV}$.

\begin{figure}
    \centering
    \includegraphics[scale=0.8]{u1blc}
    \caption{Constraints in the $\operatorname{U}(1)_{B-L}$ model.}
    \label{fig:u1blc}
\end{figure}

The constraint of $\Delta N_{\text{eff}}$ for $\operatorname{U}(1)_{B-L}$ is shown in the blue (middle) region and start to be better than current ATLAS limit for $M_{Z'}\gtrsim 4.8\ \text{TeV}$. Similar restrictions can be obtained for the other models quoted in fig.~\ref{fig:Neff}.

\section{Discussion and summary}

\appendix
\section{Anomaly cancellation}
We use $f$ ($f$) to denote the general $\operatorname{U}(1)_X$ generation-independent charge assignments of the field $f_R$ ($F_L$). The three linear anomalies in $\operatorname{U(1)_X}$~\cite{Campos:2017dgc}
\begin{align}
  \label{eq:anolin}
  \left[\mathrm{SU}(3)_{c}\right]^{2} \mathrm{U}(1)_{X} :& &
[3 u+3 d]-[3 \cdot 2 q]=&0 \nonumber\\
\left[\mathrm{SU}(2)_{L}\right]^{2} \mathrm{U}(1)_{X} :&&
-[2 l+3 \cdot 2 q]=&0 \nonumber\\
\left[\mathrm{U}(1)_{Y}\right]^{2} \mathrm{U}(1)_{X} :&&
{ \left[(-2)^{2} e+3\left(\tfrac{4}{3}\right)^{2} u+3\left(-\tfrac{2}{3}\right)^{2} d\right]-\left[2(-1)^{2} l+3 \cdot 2\left(\tfrac{1}{3}\right)^{2} q\right]}=&
 0
  \end{align}
allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $N$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \left[\text{Grav}\right]^{2} \mathrm{U}(1)_{X}:\ \sum_{\alpha=1}^N n_{\alpha}+3 (e-2l)=&0\,, &  \left[\mathrm{U}(1)_{X}\right]^{3}:\ \sum_{\alpha=1}^N n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
We choose the solutions with $\lambda\equiv e-2l$, such that
\begin{align}
  \label{eq:anolam}
  \sum_{\alpha=1}^{n} n_{\alpha}=&-3 \lambda\,,   & \sum_{\alpha=1}^{n} n_{\alpha}^3=&-3 \lambda^3\,, 
\end{align}
The full set of standard model charges in terms of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc} that we choose as $l$ and $\lambda$, is just
\begin{align}
  u=&-\lambda-\frac{4l}{3}\,,&d=&\lambda+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&\lambda+2l\,,&h=&-\lambda-l\,.
  \label{Eq:SMCharges}
\end{align}
If we now change $f\to  f/\lambda$, such that $\lambda\to 1$, the first of anomaly cancellation~\eqref{eq:anolin} remain invariant, and without lost of generality
always is possible to normalize the solutions such that the last set~\eqref{eq:anolam} is just
\begin{align}
  \label{eq:anolamnor}
   \sum_{\alpha=1}^{n} n_{\alpha}=&-3\,,   & \sum_{\alpha=1}^{n} n_{\alpha}^3=&-3\,.
\end{align}
For example, the solution with $\lambda=3$: $n_{\alpha}=\left( -2,-2,-4,-1 \right)$~\cite{Appelquist:2002mw} can be easily normalized to the form \eqref{eq:anolamnor} with $n_{\alpha}\to n_{\alpha}/3$ as used in~\cite{Patra:2016ofq}. We work with the normalized solution in terms of a single parameter~\cite{Jenkins:1987ue,Oda:2015gna,Okada:2018tgy}, that we choose to be $l$, as summarized in column $\operatorname{U}(1)_X$ of Table~\ref{tab:partcont3}. In particular, this includes the solution  $n_{\alpha}=(-4,-4,+5)$~\cite{Appelquist:2002mw}.

>>>>

\begin{abstract}    
We propose a model for one-loop Dirac neutrino masses with Majorana mediators which can be dark matter candidates if they are the lightest states circulating the loop. This model is restricted by  neutrino physics,  lepton flavor violation and cosmological constrains, including  
the effective number light neutrinos $N_{\text{eff}}$. Here we find that $M_{Z^{\prime}}/g^{\prime} \gtrsim 40\ \text{TeV}$ in order to satisfy the conditions of $N_{\text{eff}}$..
\end{abstract}

@@

\section{Beyond the minimal model: cosmological and collider constraints}
\label{sec:CosmoConstraints}
It turns that the $Z'$ portal also allows to probe the model through modifications on the cosmological history of the Universe, namely, via additional contributions to the effective number of relativistic degrees of freedom $N_{\text{eff}}$~\cite{Dolgov:2002wy}. In this model, these contributions arise from the presence of the right-handed neutrinos and may be expected to be sizeable precisely due to the large $\operatorname{U}(1)_{\mathcal{D}}$~charges of the $\nu_R$'s.  Nevertheless, since the right-handed neutrinos do not couple directly to the rest of the SM particles (see Eq.~(\ref{eq:LagY}))  they  decouple early enough from the thermal bath and, therefore, do not modify the SM prediction for $N_{\text{eff}}$.

@@

%
\begin{figure}[t]
\centering
\includegraphics[scale=1]{D_Neff.pdf}
\caption{Contribution to the number of extra relativistic degrees of freedom ($\Delta N_{\text{eff}}$) in function of $M_{Z^{\prime}}/g^{\prime}$. The region above the gray line is excluded by the measurements at $2\sigma$ from PLANCK~\cite{Aghanim:2018eyx}. The horizontal dashed lines show the projected sensitivity of the future experiments SPT-3G/SO~\cite{Benson:2014qhw} and CMB-S4~\cite{Abitbol:2019nhf}. }
\label{fig:Neff}
\end{figure}
%

The results for $\Delta N_{\text{eff}}$ as a function of $M_{Z^{\prime}}/g^{\prime}$ are displayed in Fig.~\ref{fig:Neff}. It follows that for small rartios $M_{Z'}/g'$  the variation in the number of relativistic degrees of freedom of the new species is large ($\Delta N_{\text{eff}}\gtrsim0.3$ corresponds to decoupling temperatures $T^{\nu_R}_{\text{dec}}\lesssim2$ GeV) and vice versa.
% The bound that come from Planck~\cite{Aghanim:2018eyx} is stronger than the bound of LEP~\cite{Heeck:2014zfa}. and even LHC, see below
It follows that for the $\operatorname{U}(1)_{G}$ model the Planck+BAO upper bound demands that $M_{Z^{\prime}}/g^{\prime} \gtrsim 40$ TeV while for the $\operatorname{U}(1)_{R}$ model a more stringent lower bound applies $M_{Z^{\prime}}/g^{\prime} \gtrsim 60$ TeV. 
%in order to satisfy the conditions of $N_{\text{eff}}$ for all the symmetries studied here.
On the other hand, it is remarkable the fact the future CMB experiments~\cite{Benson:2014qhw,Abitbol:2019nhf} have the potential to entirely probe all the not-so minimal models~\cite{Abazajian:2019oqj}.  
%Prospect on $N_{\text{eff}}$~\cite{Abazajian:2019oqj}

%\section{Collider constraints}
Regarding collider searches, the recasting of the latest ATLAS results for the search of dilepton resonances using $139\ \text{fb}^{-1}$~\cite{Aad:2019fac} was done
in Ref.~\cite{Chiang:2019ajm} for the $\operatorname{U}(1)_{B-L}$ model.
The green (upper) region in Fig.~\ref{fig:u1blc} shows the excluded region at $95\%\ \text{C.L.}$ To ease the comparison with other results, we show the exclusion as function of $M_{Z'}/g^{\prime}$. In particular, the limit from LEP for $\operatorname{U}(1)_{B-L}$ model is~\cite{Carena:2004xs,Heeck:2014zfa}\footnote{The constraint as function of the $X$-charge $h=-1-l$ is given in Ref.~\cite{Okada:2016tci}. }
\begin{align} 
  M_{Z'}/g^{\prime}>6.7\ \text{TeV}\,,&&\text{at $95\%$ C.L }\,,
\end{align}
which is obtained from the search for effective four-lepton operators and is valid for $M_{Z'}\gg 200\ \text{GeV}$. This constraint corresponds to the excluded magenta (lower) region of Fig.~\ref{fig:u1blc} and start to be relevant for $M_{Z'}>5.8\ \text{TeV}$.

\begin{figure}
    \centering
    \includegraphics[scale=1.0]{u1blc}
    \caption{Collider and cosmological constraints for the $\operatorname{U}(1)_{B-L}$ model.}
    \label{fig:u1blc}
\end{figure}

The constraint of $\Delta N_{\text{eff}}$ for $\operatorname{U}(1)_{B-L}$ is shown in the blue (middle) region and start to be better than current ATLAS limit for $M_{Z'}\gtrsim 4.8\ \text{TeV}$. Similar restrictions can be obtained for the other models quoted in Fig.~\ref{fig:Neff}.

\section{Summary}

\section*{Acknowledgments}
Work supported by Sostenibilidad-UdeA and the UdeA/CODI Grant 2017-16286, and by COLCIENCIAS through the Grant 111577657253. 
\appendix
\section{$\operatorname{U}(1)_X$ anomaly cancellation }
We use $f$ ($f$) to denote the general $\operatorname{U}(1)_X$ generation-independent charge assignments of the field $f_R$ ($F_L$). The three linear anomalies in $\operatorname{U(1)_X}$~\cite{Campos:2017dgc}
\begin{align}
  \label{eq:anolin}
  \left[\mathrm{SU}(3)_{C}\right]^{2} \mathrm{U}(1)_{X} :& &
[3 u+3 d]-[3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{SU}(2)_{L}\right]^{2} \mathrm{U}(1)_{X} :&&
-[2 l+3 \cdot 2 q]=&0, \nonumber\\
\left[\mathrm{U}(1)_{Y}\right]^{2} \mathrm{U}(1)_{X} :&&
{ \left[(-2)^{2} e+3\left(\tfrac{4}{3}\right)^{2} u+3\left(-\tfrac{2}{3}\right)^{2} d\right]-\left[2(-1)^{2} l+3 \cdot 2\left(\tfrac{1}{3}\right)^{2} q\right]}=&
 0,
  \end{align}
allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $N$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \left[\text{Grav}\right]^{2} \mathrm{U}(1)_{X}:\ \sum_{\alpha=1}^N n_{\alpha}+3 (e-2l)=&0\,, &  \left[\mathrm{U}(1)_{X}\right]^{3}:\ \sum_{\alpha=1}^N n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
We choose the solutions with $\lambda\equiv e-2l$, such that
\begin{align}
  \label{eq:anolam}
  \sum_{\alpha=1}^{n} n_{\alpha}=&-3 \lambda\,,   & \sum_{\alpha=1}^{n} n_{\alpha}^3=&-3 \lambda^3\,.
\end{align}
The full set of standard model charges in terms of two parameters~\cite{Appelquist:2002mw,Campos:2017dgc} that we choose as $l$ and $\lambda$, is just
\begin{align}
  u=&-\lambda-\frac{4l}{3}\,,&d=&\lambda+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&\lambda+2l\,,&h=&-\lambda-l\,.
  \label{Eq:SMCharges}
\end{align}
If we now change $f\to  f/\lambda$, such that $\lambda\to 1$, the first of anomaly cancellation~\eqref{eq:anolin} remain invariant, and without lost of generality
always is possible to normalize the solutions such that the last set~\eqref{eq:anolam} is just
\begin{align}
  \label{eq:anolamnor}
   \sum_{\alpha=1}^{n} n_{\alpha}=&-3\,,   & \sum_{\alpha=1}^{n} n_{\alpha}^3=&-3\,.
\end{align}
For example, the solution with $\lambda=3$: $n_{\alpha}=\left( -2,-2,-4,-1 \right)$~\cite{Appelquist:2002mw} can be easily normalized to the form in Eq.~\eqref{eq:anolamnor} with $n_{\alpha}\to n_{\alpha}/3$ as used in Ref.~\cite{Patra:2016ofq}. We work with the normalized solution in terms of a single parameter~\cite{Jenkins:1987ue,Oda:2015gna,Okada:2018tgy}, that we choose to be $l$, as summarized in column $\operatorname{U}(1)_X$ of Table~\ref{tab:partcont3}. In particular, this includes the solution  $n_{\alpha}=(-4,-4,+5)$~\cite{Appelquist:2002mw}.
